---
title: "Hand Localization Signals"
output: html_notebook
---

In this document all analyses can be reviewed with full output. It starts by loading (and installing) the package with code that is used for the project and paper.

```{r}
if (!require(handlocs)) {
  if (!require(devtools)) {
    install.packages(devtools)
  }
  # options(download.file.method = "libcurl")
  devtools::install_github('thartbm/handlocs', ref='main')
}
library(handlocs)
```

Lists of urls and raw data files are available from the `handlocs` package. We'll store them in the `data/` folder here:

```{r}
data(files, package='handlocs')
write.csv(files, 'data/files.csv', row.names = FALSE)
data(urls, package='handlocs')
write.csv(urls, 'data/urls.csv', row.names = FALSE)
```

We can now also download the raw data from OSF.

```{r}
handlocs::dataDownload(groups='all', sections=c('aligned','rotated'), overwrite=FALSE)
```

We now have all data used for the paper. But there is more data for people that want to explore that: the reaches made during the localization trials. Add the section `localization-reaches` to the command to also download that data.

The main thing to do first is to further clean the data. All trials have been manually inspected and have been tagged as selected or not: every csv file has a `selected` column with 1 for selected trials and 0 for not selected trials. All reach trajectories in `training` and `nocursor` files (but not the `activereach` and `passivereach` files) also have a sample indicated to correspond to the point of maximum velocity, which we use as point to get the angular reach deviations at.

```{r}
getFileReachDeviations <- function(data, column='maxvelocity', value=1, FUN=NULL) {
  
  
  df <- read.csv(data, stringsAsFactors = FALSE)
  
  if (!is.null(FUN)) {
    df <- FUN(df)
  }
  
  df <- df[which(df[[column]] == value),]
  
  for (target in unique(df$targetangle_deg)) {
    
    idx <- which(df$targetangle_deg == target)
    
    df[idx,c('handx_cm','handy_cm')] <- SMCL::rotateCoordinates(df[idx,c('handx_cm','handy_cm')], -target)
    
  }
  
  df$angulardeviation_deg <- (atan2(df$handy_cm,df$handx_cm) / pi) * 180
  
  df
  
}
```


