---
title: "Hand Localization Signals"
output:
  word_document: default
  html_notebook: default
---

In this document all analyses can be reviewed with full output. It starts by loading (and installing) the package with code that is used for the project and paper. The code in this package is not specific to this paper

```{r}
if (!require(handlocs)) {
  if (!require(devtools)) {
    install.packages(devtools)
  }
  # options(download.file.method = "libcurl")
  devtools::install_github('thartbm/handlocs', ref='main')
}
library(handlocs)
```

Lists of urls and raw data files are available from the `handlocs` package. We'll store them in the `data/` folder here:

```{r}
data(files, package='handlocs')
write.csv(files, 'data/files.csv', row.names = FALSE)
data(urls, package='handlocs')
write.csv(urls, 'data/urls.csv', row.names = FALSE)
```

We can now also download the raw data from OSF.

```{r}
handlocs::dataDownload(groups='all', sections=c('aligned','rotated'), overwrite=FALSE)
```

We now have all data used for the paper. But there is more data for people that want to explore that: the reaches made during the localization trials. Add the section `localization-reaches` to the command to also download that data.

The main thing to do first is to further clean the data. All trials have been manually inspected and have been tagged as selected or not: every csv file has a `selected` column with 1 for selected trials and 0 for not selected trials. All reach trajectories in `training` and `nocursor` files (but not the `activereach` and `passivereach` files) also have a sample indicated to correspond to the point of maximum velocity, which we use as point to get the angular reach deviations at.

```{r}
getFileReachDeviations <- function(data, column='maxvelocity', value=1, FUN=NULL) {
  
  
  df <- read.csv(data, stringsAsFactors = FALSE)
  
  if (!is.null(FUN)) {
    df <- FUN(df)
  }
  
  df <- df[which(df[[column]] == value),]
  
  for (target in unique(df$targetangle_deg)) {
    
    idx <- which(df$targetangle_deg == target)
    
    df[idx,c('handx_cm','handy_cm')] <- SMCL::rotateCoordinates(df[idx,c('handx_cm','handy_cm')], -target)
    
  }
  
  df$angulardeviation_deg <- (atan2(df$handy_cm,df$handx_cm) / pi) * 180
  
  df
  
}
```

# Hand Localization

(remove localizations that have been de-selected, are more than 4 cm from the arc, and with Y<0 - no other screening)

## Biases throughout the workspace

Summarize at 3 training targets? Effect of age?

## Efferent and Afferent precision

Effect of age? (2 scatters, with strip-plots/denisities in the margins)

# Reaching

(get reach deviations at maximum velocity in training and at endpoint in no-cursor)

## Reach accuracy with and without cursor

What is the accuracy in each case? Biases to each target. Effects of age?

## Reach precision with and without cursor

What is the precision of reaches? Does this actually change after adaptation? Effects of age?

# Localization and Reaching

## Bias

Do biases in localization (accuracy per target), predict biases in reaching? With or without cursors? Effects of age?

## Precision

Does localization precsion (afferent or efferent?) predict precision in reaching with or without cursor? Effects of age?

# Adaptation

## Localization

Does adaptation depend on localization precision (efferent / afferent)? Effects of age?

## Exploration, exploitation

Does adaptation depend on reach precision (with or without cursor)? Effects of age?

